---
title: "interactifier"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{interactifier}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(interactifier)
```

## Introduction
When doing exploratory data analysis, identifying interactions between independent variables is vital since they may have significant impacts on the outcome. For example, Qi et al. (2009) observed a significant interaction between genetic disposition and the Western dietary pattern that contributed to the epidemic of type 2 diabetes (T2D) [1]. As another example, Lee et al. (2001) found that asbestos exposure multiplied risk of lung cancer by a similar factor in non-smokers and smokers [2]. When building a regression model, we often experience difficulty in making the decision as to whether to include the interaction terms between two independent variables in our final model if the significance of the interaction is not known. Missing important interaction terms can lead to inaccurate data interpretation and poor performace of predictive models. In order to solve the dilemma, we build a packaged named "interactifier" to test for and visualize all the significant interactions between each two-order combination of the selected independent variables from a dataset, regressed against a specified outcome variable.

## Method
Our package "interactifier" is a package that enable users to test whether there are significant interactions between any of the two independent variables in a dataset and generate a 3D interactive plot between the outcome variable and any two predictor variables. This package is composed of three functions: the first one is called "interactor", which will generate a p-value vector, and each value in this vector represents the p-value of the correlation coefficient for the interaction between each of the possible combination of two independent variables. The second functin is called "corrector", which will then use the Bonferroni corrected significance level to filter out all the significant p-values. The third function in our package is called "plotter", which will generate a 3-dimensional scatter plot with the regression plane for the selected two predictor variables and the outcome variable in the dataset. With our package, users can find all the significant interactions between any of the two independent variables in the dataset and visualize specific interactions in a 3-D graph.

The first function in our package is called "interactor". "Interactor" will loop over all the possible combinations with the size of 2 variables in the whole dataset. For each iteration, interactor will take two predictors from all of the independent variables and fit a linear regression model between each two selected predictors and the outcome variable, while taking into consideration the interaction between the two selected predictors. After fitting the linear regression model, interactor will return the p-value for the interaction term. The iteration continues until all the possible combinations with the size of two variables are tested. After the iterations are completed, the interactor function will finally return a p-value vector, with each of the p-values in this vector corresponding to a two-variable interaction term. 

To call the interactor function, the user need specify three arguments: "data", "predictors", and "outcome". The first argument "data" should be a dataframe, from which the predictor variables and the outcome variable will be drawn. The second argument "predictors" is a character vector comprises the names of all predictor variables in the dataset.The third argument "outcome" should be the name of the outcome variable in the dataset.
```{r cars}
#Use this block to view the function documentation
?interactor
```

The second function in our package is called "corrector". "Corrector" will take the p-value vector generated from our first function "interactor", and compare all the p-values in this vector with the Bonferroni-corrected alpha value. If the p-value for the interaction term is significant after the comparison, these interaction terms with their p-values will be returned. The function corrector has two arguments: The first argument "p_value" is a vector comprising all the p-values for the interaction terms. The second argument alpha is the significance level with the default setting to 0.05.
```{r}
#Use this block to view the function documentation
?corrector
```

The third function in our package is called "plotter". "Plotter" will take the outcome variable and 2 predictor variables from the dataset and generate an interactive 3D scatter plot with a regression plane. The regression plane will take into consideration the interaction between the two predictor variables as well. To call the plotter function, the user need to specify four arguments: the dataframe "my_df", the first predictor "var1", the second predictor variable "var2",  and the outcome variable "outcome". Please note that the users need to use the quotation marks on the variable names.
```{r}
#Use this block to view the function documentation
?plotter
```

To use our package to filter all the signifocant interactions between any two independent variables, users need to call the function interactor first to generate the p-value vector, and then call the function corrector to find any significant p-values under the Bonferroni corrected significance level.

##Result
We used the dataset Boston, from the MASS package, to test the functions in our package. The Boston Housing Dataset is a derived from information collected by the U.S. Census Service concerning housing in the area of Boston MA. Below is a brief introduction of all the variables in the dataset:

crim: per capita crime rate by town.
zn: proportion of residential land zoned for lots over 25,000 sq.ft.
indus: proportion of non-retail business acres per town.
chas: Charles River dummy variable (= 1 if tract bounds river; 0 otherwise).
nox: nitrogen oxides concentration (parts per 10 million).
rm: average number of rooms per dwelling.
age: proportion of owner-occupied units built prior to 1940.
dis: weighted mean of distances to five Boston employment centres.
rad: index of accessibility to radial highways.
tax: full-value property-tax rate per $10,000.
ptratio: pupil-teacher ratio by town.
black: 1000(Bk - 0.63)^2 where Bk is the proportion of blacks by town.
lstat: lower status of the population (percent).

We used "medv" as the outcome variable, and all the other variables as the indepedent variables (i.e. predictors). We then tested the significant interactions between each two-order combination of the predictors in the Boston dataset. We first apply the interactor function on the Boston dataset, then specify the first argument "data" as "Boston, the second argument "predictors" as a vector containing the names of all the indepedent variables, and the third argument "outcome" as the the outcome variable "medv", and finally stored the result in a variable named "INTERACTIONPVALS". "INTERACTIONPVALS" contains is a p-value vector containing all the p-values for each of the two-variable interactions. We then apply the corrector function on the p-value vector "INTERACTIONPVALS", and store the result in the variable "SIGPVALS". "SIGPVALS" returned us a list containing all the significant interactions terms and their corresponding p-values. Below is the presentation of how we test the functions in our package.

```{r}
INTERACTIONPVALS<-interactor(data=Boston,predictors=c('indus',"crim","zn","chas","nox","rm","age","dis","rad","tax","ptratio","black","lstat"), outcome='medv')
INTERACTIONPVALS

SIGPVALS<-corrector(INTERACTIONPVALS)
SIGPVALS

plotter(Boston, var1="indus", var2="rm",outcome="medv")
```

##Conclusion
In conclusion, we successfully created a package named "interactifier" to loop over all the independent variables in a dataset to find all the significant interactions. By implementing our package, users will be able to recognize significant interactions between any two of the independent variables and also make decisions of which interaction terms they may want to include in their final model.

There are several limitations of our package: First, we only tested two-variable interactions, when higher order interaction may exist among independent variables. Second, each time we built the linear regression model, we only included two predictors, when other predictors may also play significant roles and affect the outcome variable simultaneously. Finally, the application of our package may be limited to dataset that use linear regression models for predictions. Future studies can look further into higher order interactions and extend our package to make it fit generalized linear models and non-linear models.

##References:
[1] Qi L, Cornelis MC, Zhang C, van Dam RM, Hu FB. Genetic predisposition, Western dietary pattern, and the risk of type 2 diabetes in men. Am J Clin Nutr. 2009;89(5):1453–1458. doi:10.3945/ajcn.2008.27249
[2] Lee PN. Relation between exposure to asbestos and smoking jointly and the risk of lung cancer. Occup Environ Med. 2001;58(3):145–153. doi:10.1136/oem.58.3.145
